Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-10 examples.tex
+

Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{entry}$) is present or not (Line 21 in Figure \ref{fig:callptetoinitialize}). To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}), and then,check if \textit{presence} bit of the entry is set or not. We see in Specification Line 6 that \textsf{entry} has \textsf{pte-pointsto} that is constructed on the client of \textfsf{pte\_get\_next\_table} -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir}.

@@

Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \texstsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$. This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.

@@

\end{array}
\]
\vspace{-1em}
\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
  \label{fig:peraspaceinvariant_with_p2v_extension}
\end{figure*}
\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map $\Xi$ to construct the \ref\textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltoptepteinitialize}. 
Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
of the value in memory --- so full ownership is returned only for unused entries.
This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.

>>>>

Within \textsf{get\_next\_table}, first, we check if the entry whose address is passed as a parameter ($\textsf{entry}$) is present or not (Line 21 in Figure \ref{fig:callptetoinitialize}). To be able to do so, it needs to load the entry value (a 64-bit word divided into bit-fields for corresponding fileds shown in Section {}), and then,check if \textit{presence} bit of the entry is set or not. We see in Specification Line 6 that \textsf{entry} has \textsf{pte-pointsto} that is constructed on the client of \textsf{pte\_get\_next\_table} -- e.g.\textsf{pte\_walkpgdir} in Figure \ref{walkpgdir}.

@@

Looking at the update from Specfication Lines 5-10 to 10-14 in Figure \ref{walkpgdir}, we see that the removal of \textsf{entry} from the map $\Xi$ in the invariant to construct the \textsf{pte-ptsto} (Specfication Line 13 in Figure \ref{walkpgdir}) needed for executing the load instruction (Line 15 in Figure \ref{fig:calltopteinitialize}) to obtain the value of a page entry whose physical address (\textsf{entry}) kept in the register $\texttt{r8}$. This magical abstraction -- represented as $\Xi$ in the invariant definition -- provides us to construct this \textsf{pte-ptsto} out of an physical address is called \textit{identity-mappings}.

@@

\end{array}
\]
\vspace{-1em}
\caption{Global Address-Space Invariant in Figure \ref{fig:peraspaceinvariant} extended with a ghost map bookkeeping identity mappings }
  \label{fig:peraspaceinvariant_with_p2v_extension}
\end{figure*}
\textsf{pte\_walkpgdir}, as a client, holds the knowledge that there exists an identity mapping for the physical entry address (\textsf{entry}) in the root page table ($\textsf{L}_{4}$):  $\mathsf{entry} \mapsto_{\textsf{id}} \textsf{\_}$ in Specification Line 3 is a partially owned token for accessing and looking up the resources in the identity map, $\Xi$, to construct the \textit{virtual-to-physical} pointsto relation $\textsf{entry+KERNBASE} \mapsto_{\textsf{vpte,qfrac}} \textsf{entry \entry\_val}$ with the virtual address (\textsf{entry+KERNBASE}) obtained by offsetting the physical address (\textsf{entry}). With this knowledge on the root-page-table-entry, we can start traversing the page-table tree which requires locating the address of the next table -- a call to \textsf{pte\_get\_next\_table} shown in Figure \ref{fig:calltoptepteinitialize}. 
Beyond a frame, the precondition before Line 15 requires the current address space invariant, and knowledge that \textsf{entry} is mapped to a random entry value, subtly, the operation also, at least, requires that the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
of the value in memory --- so full ownership is returned only for unused entries.
This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
